---
title: "Jumia-Web-scraping Project"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Jumia Web-scraping Project

-Using R to scrape Jumia's website products and their discounts on `r lubridate::today()`
```{r}
# loading necessary packages
library(rvest)
library(dplyr)
library(httr)

```

## web scraping
```{r}
url= "https://www.jumia.co.ke/"

# reading url
page = read_html(url)

# checking response
(httr::status_code(GET(url))) # its 200
```

```{r}
# Extracting all category tabs that contain all the products 
divs = page %>% html_nodes("div[class='col16 -pvs']") #saving to div variable

# creating an empty data frame to save the data
jumia = tibble(  "Name"= character(),      # name column
         "Price" =character(),             # price column
         "Discount" =character(),          # discount column
         "Category"= character())          # category column
```


```{r}
# creating a function that extracts all the products information in a certain category
category = function(div){ # passing a div parameter
  
  # selecting all items in that div
  items = div  %>% html_nodes("div[class='itm col']") 
  
  # assigning category as the div's title/header in the website
  category = div  %>%html_nodes("h2") %>% html_text()  # text in the h2 tag
  
  name = items %>% html_node("div[class='name']") %>% html_text() # all item names
  price = items %>% html_node("div[class='prc']") %>% html_text() # all item prices
  discount = items %>% html_node("div[class='bdg _dsct']") %>% html_text() #all item discounts
  
 # returning a data frame of the extracted items
 return(tibble("Name"= name,  
         "Price" = price,
         "Discount" = discount,
         "Category"=  category) )

}

# running the above category function to all the category divs and appending the data to the jumia data frame
for(div in divs[c(2,4:23)]){
  jumia= bind_rows(jumia,category(div))
}

# previewing our data
head(jumia, n=10)
```

## Cleaning and saving the scraped data
```{r}
# converting the price and discount columns to integers

# In both columns we will replace all non numeric numbers with a blank
jumia = jumia %>% 
  mutate(Price = as.numeric(Price %>%
                              stringr::str_replace_all(pattern="[^0-9]","") )
         ,
         Discount= as.numeric(Discount %>% 
                                stringr::str_replace_all(pattern="[^0-9]","")  ))

# previewing clean data
head(jumia)
```


```{r}
# saving as csv
readr::write_csv(jumia,"../data/jumia_data.csv")
```


# Data Analysis
```{r}
# loading required packages
library(dplyr)
library(ggplot2)

```

- Bellow is the list of the scarped categories and the count of items in each.
```{r}

jumia %>% 
  group_by(Category) %>% count() %>% arrange(desc(n))

```

-Below is a barplot of the various categories and their average discounts on products
```{r}
# bar plot of categories vs discount

# creating plot
plot= jumia %>% 
  group_by(Category) %>% 
  summarize(count=n(),discount= mean(Discount, na.rm = T)) %>% 
  arrange(desc(discount)) %>% 
  ggplot(aes(x=forcats::fct_reorder(Category,desc(discount)), 
               y=discount))+
  geom_col() + # bar plot layer
  geom_label(aes(label = paste(round(discount,1),"%")),size=3) #adding discount as text labels

# formating the plot
plot+
  # adding x and y labels
  labs(title = "Barplot of Categories Vs Discount",
       y="Discount",
       x="Categories")+
  theme_bw()+
  # rotating the x labels
  theme(axis.text.x = element_text(angle = 90))
  

```
-Clearance Deals have the highest discounted products, as the category has a discount of 43.7%, hence if in need of large discounts go to the clearance deals category followed by Men's shoes and Deals on Phones and accessories.


-Below is a bar plot of the top 5 categories with the lowest prices
```{r}
# barp lot of categories vs Price

# creating plot
plot= jumia %>% 
  group_by(Category) %>% 
  summarize(count=n(),price= mean(Price, na.rm = T)) %>% 
  filter( rank(price)<=5) %>% 
  ggplot(aes(x=forcats::fct_reorder(Category,price), 
               y=price))+
  geom_col() + # bar plot layer
  geom_label(aes(label = round(price,1)),size=3) #adding discount as text labels

# formatting the plot
plot+
  # adding x and y labels
  labs(title = "Top 5 Categories with Lowest Prices",
       y="Price",
       x="Categories")+
  theme_bw()+
  # rotating the x labels
  theme(axis.text.x = element_text(angle = 90))
  
```
-Body shop has products with the lowest prices followed by Clearance Deals and Drink up discounts

-Bellow ae the top 10 products with the highest discounts today
```{r}
# filtering the data
data = jumia %>% 
  arrange(desc(Discount)) %>% 
  filter(rank(desc(Discount))<=5) %>% 
  mutate(Name=stringr::str_sub(Name,1,16))

  # creating the plot
ggplot(data=data, aes(x= forcats::fct_reorder(Name,Discount),
                      y=Discount, 
                      label=paste("Disc:",Discount," Price:",Price)))+
  geom_col(inherit.aes = T)+
  geom_label(size=3)+
  labs(title = "Top 10 Products with highets Discounts",
       y="Discount",
       x="Product Name")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90))
  
```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```

